---
title: "Covid-19 lockdown measures and GDP growth"
output: html_notebook
---

# Housekeeping & packages

```{r}
rm(list = ls())
cat("\014")
options(scipen = 999)
library(bundesbank)
library(rdbnomics)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(readr) # parse_double
library(ggrepel)
library(forcats)
library(wbstats)
library(openxlsx)
```

# GDP growth in H1

## DBnomics

### Eurostat & OECD

Construct codes for DBnomics
```{r}
codes_eurostat <- c("DE",
                    "BE",
                    "FR",
                    "IT",
                    "ES",
                    "IE",
                    "AT",
                    "DK",
                    "EL",
                    "FI",
                    "HU",
                    "CZ",
                    "PL",
                    "PT",
                    "SE",
                    "NL",
                    "SI")
codes_oecd <- c("AUS",
                "BRA",
                "CAN",
                "CHL",
                "COL",
                "IND",
                "IDN",
                "ISR",
                "JPN",
                "KOR",
                "MEX",
                "GBR",
                "USA",
                "TUR",
                "CHE")

codes_dbnomics <- c(paste0("Eurostat/teina011/Q.B1GQ.PCH_Q1_SCA.", codes_eurostat),
                    paste0("OECD/MEI/", codes_oecd, ".NAEXKP01.GPSA.Q"))
```

Call `rdb`
```{r}
df <- rdb(provider_code = codes_dbnomics) %>%
        filter(original_period %in% c("2020-Q1", "2020-Q2")) %>%
        mutate(original_value = parse_double(original_value),
               country = ifelse(is.na(Country), `Geopolitical entity (reporting)`, Country)) %>%
        select(original_value, original_period, country) %>%
        group_by(country) %>%
        summarize(gdpH1 = 100 * (1+original_value[original_period == "2020-Q1"]/100) * 
                                (1+original_value[original_period == "2020-Q2"]/100) - 100)

df$country[grepl(pattern = "Germany", df$country)] <- "Germany"
```

## Bundesbank


```{r}
codes_Bbk <- c("BBXN1.Q.CN.S.NAG.B1QG00.0000.Z0N.PER.I00",
               "BBXN1.Q.RU00.S.NAG.B1QG00.0000.Z0N.CPC.I00")

countries_Bbk <- c("China", "Russia")

for (i in seq_along(codes_Bbk))
{
  if (countries_Bbk[i] == "China") # China already in q/q growth rate!
  {
      tmp <- getSeries(codes_Bbk[i], start = "2020-01", end = "2020-06")


      gdpH1 <- 100 * (1+tmp[1, 2]/100) * (1+tmp[2, 2]/100) - 100
      df <- rbind(df, data.frame(country = countries_Bbk[i],
                                 gdpH1 = gdpH1))

}
  else
  {
      tmp <- getSeries(codes_Bbk[i], start = "2019-09", end = "2020-06")
      #tmp$values <- parse_double(tmp$values)
      gdpH1 <- tmp[3, 2] / tmp[1, 2] * 100 - 100
      df <- rbind(df, data.frame(country = countries_Bbk[i],
                                 gdpH1 = gdpH1))
  }
}
```



# Stringency index

## Load raw data and select the relevant indicators

```{r}
rawdata_string <- read_csv("https://github.com/OxCGRT/covid-policy-tracker/raw/master/data/OxCGRT_latest.csv")


df_string <- rawdata_string %>% 
             mutate(Date = ymd(Date),
                    CountryName = CountryName,
                    StringencyIndex = StringencyIndex) %>% 
             select(CountryName, StringencyIndex, Date) 
```

## Calculate average index value for 2020H1

```{r}
df_string_avgH1 <- df_string %>% 
                   filter(Date <= "2020-06-30") %>%
                   group_by(CountryName) %>%
                   summarise(stringency = mean(StringencyIndex))
```

Merge with GDP growth in 2020H1
```{r}
df <- merge(df, df_string_avgH1, by.x = "country", by.y = "CountryName")
```


# October 2019 WEO

```{r}
GetWEOData <- function(filename, year, var, rename_col)
{
  require(dplyr)
  require(readr)
  tmp <- read_delim(filename, delim = ";") %>%
         select(country = Country, iso = ISO, variable = "WEO Subject Code", val = !!year) %>%
         filter(variable == !!var) %>%
         select(-variable, everything()) %>%
         filter(val != "n/a") %>%
         mutate(val = parse_double(val)) %>%
         select(country, iso, !!rename_col := val)
  return(tmp)
}

filename <- "WEOOct2019all.csv"

weo <- GetWEOData(filename, year = "2020", var = "NGDP_RPCH", rename_col = "gdp_Oct19WEO")

```

Rename Korea, drop iso

```{r}
weo$country[weo$country == "Korea"] <- "South Korea"

weo <- select(weo, -iso)
```

Recalculate growth rate for 2020 to correspond to 2020H1 (i.e. de-annualize)

```{r}
weo$gdp_Oct19WEO  <- (1 + weo$gdp_Oct19WEO / 100) ^ 0.5 * 100 - 100
```

Merge with `df`

```{r}
df <- merge(df, weo, by = "country")
```

Create new variable: deviation from IMF WEO forecast and remove WEO forecast
```{r}
df <- df %>% mutate(dev_gdp = gdpH1 - gdp_Oct19WEO) %>% 
             select(-gdp_Oct19WEO)
```

# Openness

Sum of exports and imports in relation to GDP. Source: IMF International Financial Statistics


```{r}
start.time <- Sys.time()
```

```{r}
mnemonics_IFS <- c("DE",
                    "BE",
                    "FR",
                    "IT",
                    "ES",
                    "IE",
                    "AT",
                    "DK",
                    "GR",
                    "FI",
                    "HU",
                    "CZ",
                    "PL",
                    "PT",
                    "SE",
                    "NL",
                    "SI",
                    "GB",
                    "US",
                    "JP",
                    "KR", 
                    "AU",
                    "CH",
                    "CA",
                    "IL",
                    "CN",
                    "RU",
                    "IN",
                    "ID",
                    "TR",
                    "BR",
                    "CL",
                    "CO",
                    "MX")

vars <- c("NGDP", "NX", "NM")
tmp1 <- rep(paste0("IMF/IFS/A.", mnemonics_IFS), each = length(vars))
tmp2 <- rep(paste0(vars, "_XDC"), times = length(tmp1) / length(vars))
codes <- paste0(tmp1, ".", tmp2)
  ifs <- rdb(provider_code = codes) %>%
         select(original_period, original_value, country = `Reference Area`, INDICATOR) %>%
         filter(original_period == 2018) %>% 
         mutate(original_value = as.numeric(original_value)) %>%
         group_by(country) %>%
         summarise(openness = (original_value[INDICATOR =="NX_XDC"] + 
                               original_value[INDICATOR =="NM_XDC"]) / 
                               original_value[INDICATOR == "NGDP_XDC"] * 100
                   )
```

Merge with `df`

```{r}
df <- merge(df, ifs, by = "country")
```

# Tourism 

As a share of GDP. Source: World Bank (WDI?)

```{r}
wb_raw <- wb_data(country = "all", indicator = c("ST.INT.RCPT.CD", "NY.GDP.MKTP.CD"), start_date = 2018)
mnemonics_wb <- mnemonics_IFS
wb_raw %>% filter(date == "2018", iso2c %in% mnemonics_wb) %>% select(country, receipts = ST.INT.RCPT.CD, gdpusd = NY.GDP.MKTP.CD, iso3c) %>%
mutate(tourism = receipts / gdpusd * 100) %>%
select(-c(receipts, gdpusd)) -> wb
wb$country[wb$country == "Russian Federation"] <- "Russia"
wb$country[wb$country == "Korea, Rep."] <- "South Korea"

```

Merge with `df` 
```{r}
df <- merge(df, wb, by = "country")
```


Alternative source: WTTC, data for travel and tourism value added. 

```{r}
wttc_raw <- read.csv(file = "WTTC.csv", sep = ";", stringsAsFactors = FALSE)

wttc_raw$Country[wttc_raw$Country == "Russian Federation"] <- "Russia"

wttc_raw %>% filter(Country %in% df$country) %>% select(country = Country, traveltourism = X2018) -> wttc
```

Merge with `df` 
```{r}
df <- merge(df, wttc, by = "country")
```


# Covid-19 deaths

Cumulated deaths by end of 2020Q2. Source: ECDC

```{r}
ecdc_raw <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", 
                    na = "")

# replace names
ecdc_raw$countriesAndTerritories[grep(pattern = "Czech+", ecdc_raw$countriesAndTerritories)] <- "Czech Republic"
ecdc_raw$countriesAndTerritories[ecdc_raw$countriesAndTerritories == "United_States_of_America"] <- "United States"
ecdc_raw$countriesAndTerritories[ecdc_raw$countriesAndTerritories == "United_Kingdom"] <- "United Kingdom"
ecdc_raw$countriesAndTerritories[ecdc_raw$countriesAndTerritories == "South_Korea"] <- "South Korea"

# calculate cumulated deaths per million inhabitants
ecdc_raw %>% mutate(date = make_date(year, month, day),
                    pop_mio = popData2019 / 1000000) %>% 
             select(country = countriesAndTerritories, geoId, date = date, cases, deaths, pop_mio) %>% 
             filter(geoId %in% mnemonics_IFS) %>%
            arrange(date) %>%
            group_by(country) %>%
            mutate(cases_cumsum = cumsum(cases) / pop_mio,
                   deaths_cumsum = cumsum(deaths) / pop_mio) %>% 
            filter(date == "2020-06-30") %>% 
            select(country, covid19cases = cases_cumsum, covid19deaths = deaths_cumsum) -> ecdc
```

Merge with `df`

```{r}
df <- merge(df, ecdc, by= "country")
```


# Google mobility data

```{r}
google_raw <- read_csv(file ="https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv")

google_raw$country_region[grep(pattern = "Czech+", google_raw$country_region)] <- "Czech Republic"

google_raw %>% filter(is.na(sub_region_1)) %>% 
               select(date, 
                      country = country_region,
                      retail_rec = retail_and_recreation_percent_change_from_baseline,
                      transit = transit_stations_percent_change_from_baseline,
                      workplace = workplaces_percent_change_from_baseline) %>%
               filter(country %in% df$country, date <= "2020-06-30") %>%
               group_by(country) %>%
               summarise(retail_rec = mean(retail_rec, na.rm = TRUE),
                         transit = mean(transit, na.rm = TRUE),
                         workplace = mean(workplace, na.rm = TRUE)) %>%
               mutate(retail_rec = retail_rec + 100,
                      transit = transit + 100,
                      workplace = workplace + 100,
                      mobility = (retail_rec + transit + workplace) / 3) -> google
```

Merge with `df` and gather into `var` and `value`

```{r}
df <- merge(df, google, by = "country", all.x = TRUE)
```

# Economic structure

GVA by sectors, source: OECD

```{r}
oecd_sect_raw <- read.xlsx(xlsxFile = "./IWH1.xlsx", sheet = "OECD Data")

oecd_sect_raw %>% select(codes, G.I = `G-I`, R.U = `R-U`) %>% mutate(structure = G.I + R.U) -> oecd_sect
```

```{r}
merge(df, oecd_sect, by.x = "iso3c", by.y = "codes") %>% select(-iso3c) %>% gather(var, value, -c(country, gdpH1, dev_gdp)) -> df
```

# Plot

List of emerging market economies
```{r}
list_EME <- c("Brazil", "Mexico", "Turkey", 
              "China", "India", "Indonesia", 
              "Chile", "Colombia", "Turkey",
              "Poland", "Hungary", "Czech Republic")
```

Calculate R^2 using function `calcR2.R`

```{r}
source("calcR2.R")
df_R2 <- calcR2(filter(df, country != "China"))
```

## Plot GDP in 2020H1 vs various indicators

`df_plot`

```{r}
df %>% select(-c(dev_gdp)) %>% 
       filter(var %in% c("stringency", "openness", "traveltourism","tourism", "covid19cases", "covid19deaths", "structure", "retail_rec"),
              country != "China") %>% 
       mutate(status = ifelse(country %in% list_EME, "Schwellenländer", "fortgeschrittene Volkswirtschaften"),
              var = fct_recode(var, "Lockdown-Stringenz" = "stringency", 
                                              "Offenheit" = "openness", 
                                              "Tourismus (WTTC)" = "traveltourism",
                                              "Tourismus (Weltbank)" = "tourism",
                                              "Covid-19-Infektionen" = "covid19cases",
                                              "Covid-19-Todesfälle" = "covid19deaths",
                                              "Mobilitätsdaten" = "retail_rec",
                                              "Dienstleistungsbereiche" = "structure")
            ) -> df_plot
```


caption text

```{r}
txt_caption <- paste0("Daten für ", length(unique(df_plot$country)), " Länder, BIP: Veränderung zwischen 2020Q2 und 2019Q4, Infektions- und Todesfälle: kumuliert, je 1 Mio. Einwohner, Stand 30.6.2020,Offenheit: Summe der Exporte und Importe im Verhältnis zur Wirtschaftsleistung,\n Tourismus (Weltbank): Einnahmen im Verhältnis zur Wirtschaftsleistung,Tourismus (WTTC): Gesamtbeitrag zur Bruttowertschöpfung (Prozent),Lockdown-Stringenz: Index, Durchschnitt in 2020H1, höhere Werte = strengere Maßnahmen,\n Dienstleistungsbereiche: Wertschöpfungsanteile der Bereiche Transport, Handel, Gast- und Freizeitgewerbe, Mobiltitätsdaten: Durchschnitt der Bereiche \"retail and recreation\", \"transit stations\" und \"workplace\", schwarze Linie: linearer Trend. \nQuelle: OECD Main Economic Indicators, Deutsche Bundesbank, ECDC, IMF International Financial Statistics, Hale et al. (2020), World Travel and Trade Council, Weltbank World Development Indicators, Google Community Mobility Reports.")
```

```{r}
df_plot$status[df_plot$country == "Germany"] <- "Deutschland"
       ggplot(df_plot, aes(x = value, y = gdpH1))+
         geom_point()+
         geom_text(data = filter(df_R2, var %in% c("Lockdown-Stringenz", "Offenheit", "Tourismus (WTTC)", "Tourismus (Weltbank)", "Covid-19-Infektionen", "Covid-19-Todesfälle", "Dienstleistungsbereiche", "Mobilitätsdaten")),
                  mapping = aes(x = x, y = y, label = R2), size = 3)+
         geom_smooth(method = "lm", se = FALSE, color = "black", size = 0.5)+
          geom_point(aes(color = status), size = 2)+
         scale_color_manual(values=c("#000000", "#00BFC4", "#FA8072"))+
         facet_wrap(~var, scales = "free_x", nrow = 2)+
         labs(y = "BIP, Prozent", x = "",
              title = "Covid-19-Pandemie: wirtschaftliche Entwicklung in 2020H1",
              subtitle = "Zusammenhang mit Covid-19, Grad der Offenheit, Lockdown-Stringenz, Wirtschaftsstruktur, Bedeutung des Tourismus und Mobilitätsdaten",
              caption = txt_caption)+
        theme_minimal()+
        theme(legend.title = element_blank(), legend.position = "top", legend.text=element_text(size=9),
              plot.caption = element_text(size = 6),
              plot.subtitle = element_text(size = 10))
```

Export to pdf

```{r}
ggsave(
  filename = "Abb_BIPH1.pdf",
  plot = last_plot(),
  device = "pdf",
  path = NULL,
  scale = 1,
  width = 25,
  height = 20,
  units = c("cm"),
  dpi = 300,
  limitsize = TRUE,
)
```




